A seamless, client-centric programming model for type safe web applications

### 2014-03-02 paper.tex
Add listing of core functions.

In order to make an RPC call using a identifier obtained from \lstinline!export!,
we must supply it with an argument. This is done using the \lstinline!<.>!
operator. It might be interesting to note that its type,
\linebreak
\lstinline!Serialize a => Export (a -> b) -> a -> Export b!,
is very similar to the type of the \lstinline!<*>! operator over applicative
functors. This is not a coincidence; \lstinline!<.>! performs the same role
for the \lstinline!Export! type as \lstinline!<*>! performs for applicative
functors. The reason for using a separate operator for this instead of making
\lstinline!Export! an instance of \lstinline!Applicative! is that since
functions embedded in the \lstinline!Export! type exist only to be called over
a network, such functions must only be applied to arguments which can be
serialized and sent over a network connection. When an \lstinline!Export!
function is applied to an argument using \lstinline!<.>!, the argument is
serialized and stored inside the resulting \lstinline!Export! object, awaiting
dispatch. \lstinline!Export! values can thus be seen as explicit
representations of closures.

@@

@@

As this modification only affects the program's entry point, which is already
given ``special'' treatment in Haskell compilers, and may even be implemented
using the plugin mechanism of GHC, requiring no modification of the compiler
itself, we believe this to be a practical, non-intrusive way to remove this
limitation, and are planning on releasing an implementation which includes
this limitation.

>>>>

In order to make an RPC call using an identifier obtained from
\lstinline!export!, we must supply it with an argument. This is done using the
\lstinline!<.>! operator. It might be interesting to note that its type,
\linebreak
\lstinline!Serialize a => Export (a -> b) -> a -> Export b!,
is very similar to the type of the \lstinline!<*>! operator over applicative
functors. This is not a coincidence; \lstinline!<.>! performs the same role
for the \lstinline!Export! type as \lstinline!<*>! performs for applicative
functors. The reason for using a separate operator for this instead of making
\lstinline!Export! an instance of \lstinline!Applicative! is that since
functions embedded in the \lstinline!Export! type exist only to be called over
a network, such functions must only be applied to arguments which can be
serialized and sent over a network connection. When an \lstinline!Export!
function is applied to an argument using \lstinline!<.>!, the argument is
serialized and stored inside the resulting \lstinline!Export! object, awaiting
dispatch. \lstinline!Export! values can thus be seen as explicit
representations of closures.

@@

A complete listing of the core functions in Haste.App is given in table
\ref{tbl:api}, and their types are given in figure \ref{lst:api}.

\begin{listingfloat}
\begin{code}
runApp       :: AppCfg -> App Done -> IO ()

runClient    :: Client () -> App Done
liftServerIO :: IO a -> App (Useless a)
export       :: Exportable a
             => a -> App (Export a)

onServer     :: Export (Server a) -> Client a
(<.>)        :: Serialize a
             => Export (a -> b) -> a -> Export b

mkUseful     :: Useless a -> Server a
getSessionID :: Server SessionID
\end{code}
\caption{Types of the Haste.App core functions}
\label{lst:api}
\end{listingfloat}

\begin{table}
\renewcommand{\arraystretch}{1.5}
\begin{center}
\begin{tabular}{|r|l|}
\hline
Function & Purpose \\
\hline
\lstinline!runApp! & Begin execution of a Haste.App program. \\
\lstinline!runClient! & \parbox[t]{5cm}{Lift a single \lstinline!Client!
computation into the \lstinline!App! monad. Must be at the very end of
an \lstinline!App! computation, which is enforced by the type system.} \\
\lstinline!liftServerIO! & \parbox[t]{5cm}{Lift an IO computation into the
\lstinline!App! monad. The computation and its result are exclusive to the
server, as enforced by the type system, and are not observable on the client.} \\
\lstinline!export! & \parbox[t]{5cm}{Make a server-side function available to
be called remotely by the client.} \\
\lstinline!onServer! & \parbox[t]{5cm}{Dispatch a remote call to the server and
wait for its completion. The result of the remote computation is returned on
the client after it completes.} \\
\lstinline!<.>! & \parbox[t]{5cm}{Apply an \lstinline!export!ed function to
a serializable argument.} \\
\lstinline!mkUseful! & \parbox[t]{5cm}{Extract the contents of a
\lstinline!Useless! value produced by \lstinline!liftServerIO! on the server.} \\
\lstinline!getSessionID! & \parbox[t]{5cm}{Get the unique identifier for
the current session. This is a pure convenience function, to relieve
programmers of the burden of session bookkeeping.} \\
\hline
\end{tabular}
\end{center}
\caption{Core functions of Haste.App}
\label{tbl:api}
\end{table}

@@

As this modification only affects the program's entry, point which is already
given ``special'' treatment in Haskell compilers, and may even be implemented
using the plugin mechanism of GHC, requiring no modification of the compiler
itself, we believe this to be a practical, non-intrusive way to remove this
limitation, and are planning on releasing an implementation which includes
this limitation.
